### 2017-04-24 Manuscript/EngineBalance.tex
Latest update. Nothing was checked, but I supposedly finished the data processing for the main engines

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{verbatim}

%opening
\title{Demonstration of the solution of the system of mass and energy balances for MS Birka Main Engines}
\author{Francesco Baldi}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Initial system}

\subsection{Mass balances}

Mass balance on the split after the compressor, with one flow going to the engine inlet valves and the other flow going to the bypass valve
\begin{equation}
\dot{m}_{air,comp} = \dot{m}_{air,cyl} + \dot{m}_{air,bypass} \\
\end{equation}

Mass balance on the mixer before the turbine, with one flow coming from the engine exhaust valves and the other flow coming from the bypass valve
\begin{equation}
\dot{m}_{eg,turb} = \dot{m}_{eg,cyl} + \dot{m}_{air,bypass} \\
\end{equation}

Note that the exhaust gas mass flow from the cylinders is fully defined as follows:
\begin{equation}
\dot{m}_{eg,cyl} = \dot{m}_{air,cyl} + \dot{m}_{fuel,cyl} \\
\end{equation}

\subsection{Energy balances}

Energy balance on the turbocharger: the power generated by the turbine must be equal to the power absorbed by the compressor
\begin{equation}
\dot{m}_{air,comp} \Delta h_{comp} = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}
\end{equation}

Energy balance on the mixer between the air flow after the bypass valve and the exhaust gas flow after the exhaust valve
\begin{equation}
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) = \dot{m}_{eg,turb} c_{p,eg} (T_{turb,in} - T_0)
\end{equation}

The system of four equations is to be solved in the four following unknowns:
\begin{itemize}
	\item $ \dot{m}_{eg,turb} $
	\item $ \dot{m}_{air,bypass} $
	\item $ \dot{m}_{air,comp} $
	\item $ T_{turb,in} $
\end{itemize}

\section{Calculating the explicit system}

First, we make all equations explicit in $ \dot{m}_{air,bypass} $ thus eliminating both 
$ \dot{m}_{eg,turb} $ and $ \dot{m}_{air,comp} $

$$
(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech} 
$$
$$
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  = (\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_0)
$$

At this point, we use the first equation to simplify the system and make it explicit in the $ \dot{m}_{air,bypass} $ variable alone:

\begin{eqnarray*}
(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,in} - T_{turb,out}) \eta_{mech}  & = & (\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) \Delta h_{comp} \\
T_{turb,in} - T_{turb,out} & = & \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} \\
T_{turb,in} & = & T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}  \\
\end{eqnarray*}

We can now substitute this expression in the second equation:

\begin{multline}
	\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
	(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} (T_{turb,out} + \frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}} \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0)
\end{multline}

We now have to simplify this equation in order to make the term $ \dot{m}_{air,bypass} $ explicit. 
\begin{comment}
One first thing we can do is noticing that the fraction:
$$
\frac{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}{\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}}
$$
Can be rewritten as:
$$
\frac{1}{1 + \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}}
$$
Since the numerator of the "subfraction" is much smaller than the denominator, we can make the assumption that this fraction behaves as:
$$
\frac{1}{1 + \epsilon}
$$
For which we know that:
$$
\frac{1}{1 + \epsilon} \approx 1 - \epsilon 
$$
That corresponds to:
$$
1 - \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}}
$$
\begin{multline}
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\ 
(\dot{m}_{eg,cyl} + \dot{m}_{air,bypass}) c_{p,eg} \left[T_{turb,out} + \left( 1 - \frac{\dot{m}_{fuel,cyl}}{\dot{m}_{air,cyl} + \dot{m}_{air,bypass}} \right) \frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}} - T_0\right]
\end{multline}
\end{comment}

We first execute the multiplication in the term on the right of the equal:
\begin{multline}
\dot{m}_{air,bypass} c_{p,air} (T_{comp,out} - T_0) + \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0)  =  \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) + \\
\dot{m}_{air,bypass} c_{p,eg} (T_{turb,out} - T_0) + \\ 
\frac{\Delta h_{comp}}{c_{p,eg} \eta_{mech}}(\dot{m}_{air,cyl} + \dot{m}_{air,bypass}) c_{p,eg}
\end{multline}

Now, we can group together the elements in the variable of interest:
\begin{multline}
\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_0) - \dot{m}_{eg,cyl} c_{p,eg}  (T_{cyl,out} - T_0) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
\end{multline}

This can be further simplified into:
\begin{multline}
\dot{m}_{air,bypass} \left[ c_{p,air} (T_{comp,out} - T_0) - c_{p,eg} (T_{turb,out} - T_0) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
\end{multline}

Finally, we can assume the specific heat at constant pressure of air and exhaust gas to be sufficiently similar for considering them equal in the term at the numerator (we ignored changes in the properties of the mixture in other points as well):
\begin{multline}
\dot{m}_{air,bypass} \left[ c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}} \right] = \\
\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}
\end{multline}

From which we can derive the final form:
\begin{equation}
\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{turb,out} - T_{cyl,out}) + \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{comp,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}} 
\end{equation}

Which can be rewritten, in order to have both numerator and denominator positive, as follows:
\begin{equation}
\dot{m}_{air,bypass} = \frac{\dot{m}_{eg,cyl} c_{p,eg} (T_{cyl,out} - T_{turb,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}\dot{m}_{air,cyl}}{c_{p,eg} (T_{turb,out} - T_{comp,out}) - \frac{\Delta h_{comp}}{\eta_{mech}}} 
\end{equation}

\end{document}